Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(ggplot2)
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot() + 
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) + 
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)

suppressMessages(library(dplyr))

pf.fc_by_age_gender <- pf %>%
  group_by(age, gender) %>%
  filter(!is.na(gender)) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n=n()) %>%
  ungroup(gender) %>%
  arrange(age, gender)

head(pf.fc_by_age_gender, 20)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = median_friend_count),
       data = pf.fc_by_age_gender) + 
  geom_line(aes(color = gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, age ~ gender, value.var = 'median_friend_count')

head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(aes(x = age, y = female / male),
       data = pf.fc_by_age_gender.wide) + 
  geom_line() + 
  geom_hline(yintercept = 1, alpha = 0.3, linetype = 2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure / 365)
head(pf$year_joined, 20)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
summary(pf$year_joined)

table(pf$year_joined)
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004, 2009, 2011, 2012, 2014))

table(pf$year_joined.bucket)


```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
table(pf$year_joined.bucket, useNA = 'ifany')

ggplot(aes(x = age, y = friend_count), data = subset(pf,!is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), 
            stat = 'summary', 
            fun.y = mean)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count), data = subset(pf,!is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), 
            stat = 'summary', 
            fun.y = mean) + 
  geom_line(stat = 'summary', 
            fun.y = mean, linetype = 2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}

friend_rate <- with(subset(pf, tenure != 0), friend_count / tenure)


median(friend_rate)
max(friend_rate)

```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated / tenure), data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket), 
            stat = 'summary', 
            fun.y = mean)

ggplot(aes(x = tenure, y = friendships_initiated / tenure), data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))


```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('./yogurt.csv')
yo$id <- factor(yo$id)

ggplot(aes(x = price), data = yo) + geom_histogram(binwidth = 10)
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
summary(yo)
length(unique(yo$price))
table(yo$price)
yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(aes(x = time, y = price), data = yo) +
  geom_point(alpha = 0.5, position = position_jitter(h=0), color = 'orange') 
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(22)
sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x = time, y = price),
       data = subset(yo, id %in% sample.ids)) + 
  facet_wrap((~id)) + 
  geom_line() + 
  geom_point(aes(size = all.purchases), pch = 1)
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r}
library(GGally)
theme_set(theme_minimal(20))
suppressMessages(ggpairs)

set.seed(1836)
pf_subset <- pf[,c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000),], axisLabels = 'internal')
```

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("green", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

### Problem Sets

```{r}
ggplot(aes(x = price, fill = cut), data = diamonds) + 
  geom_bar(binwidth = 250) + 
  facet_wrap(~color) + 
  scale_fill_brewer(type = 'qual')
```


```{r}
ggplot(aes(x = table, y = price), data = diamonds) +
  geom_point(aes(colour=cut)) +
  scale_fill_brewer(type = 'qual') #+
  #xlim(0, quantile(diamonds$table, 0.99)) + 
  #ylim(0, quantile(diamonds$price,0.99))
```

```{r}
diamonds$volume <- with(diamonds, x*y*z)
ggplot(aes(x = volume, y = price), data = diamonds) +
  geom_point(aes(colour=clarity)) + 
  coord_trans(y = "log10") +
  scale_color_brewer(type = 'div') +
  xlim(0, quantile(diamonds$volume, 0.99))
```

```{r}
pf$prop_initiated <- with(pf, friendships_initiated / friend_count)
```

```{r}
ggplot(aes(x = tenure, y = prop_initiated),
       data = subset(pf, !is.na(prop_initiated))) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = median)

ggplot(aes(x = tenure, y = prop_initiated),
       data = subset(pf, !is.na(prop_initiated))) +
  geom_smooth(aes(color = year_joined.bucket))

ggplot(aes(x = 30 * round(tenure/30), y = prop_initiated),
       data = subset(pf, !is.na(prop_initiated))) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = median)
```
```{r}
friendshipsInitiated_by_yearjoined <- group_by(subset(pf, !is.na(prop_initiated)), year_joined.bucket)
friendshipsInitiated_by_yearjoined <- summarise(friendshipsInitiated_by_yearjoined, mean_prop = mean(prop_initiated))

# Creating the graph
ggplot(aes(x = year_joined.bucket, y = mean_prop), data = friendshipsInitiated_by_yearjoined) +
  geom_bar(stat="identity") + 
  geom_text(aes(label=mean_prop), vjust=0)
```

```{r}
diamonds$pricePerCarat <- with(diamonds, price / carat)

ggplot(aes(x = cut, y = pricePerCarat), data = subset(diamonds, !is.na(pricePerCarat))) +
  geom_point(aes(colour=color), position = 'jitter') + 
  scale_color_brewer(type = 'div') +
  facet_wrap((~clarity)) 
```
```{r}
library(dplyr)
library(tidyr)
subData <- subset(gdpDataClean, year >= 1960)
subData$year <- as.numeric(subData$year)

subData$year.bucket <- cut(subData$year, breaks = c(1960, 1970, 1980, 1990, 2000, 2010, 2020))


p1<- ggplot(aes(x = year, y = value), data = subData) +
  geom_point(aes(colour=year.bucket)) + 
  #coord_trans(y = "log10") +
  scale_color_brewer(type = 'div')

p2 <- ggplot(aes(x = year, y = value),
       data = subData) +
  geom_line(aes(color = year.bucket),
            stat = 'summary',
            fun.y = mean)

p3 <- ggplot(aes(x = year, y = value),
       data = subData) +
  geom_line(aes(color = year.bucket),
            stat = 'summary',
            fun.y = median)

library(gridExtra)
grid.arrange(p1, p2, p3, ncol = 1)

ggplot(aes(x = year, y = value),
       data = subData) +
  geom_smooth(aes(color = year.bucket))

```

